---
title: "Results from spatial simulations"
author: "David L Miller"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

```{r load-pkg}
library(ggplot2)
library(reshape2)
library(plyr)
```

# Simulation setup

- No group size issues -- everything is group size 1
- No issues with $\hat{g}(0)$
- Ignoring the possibility of repeat surveys


## Detection functions

## Spatial models

- Isotropic smooths
  - Thin plate spline (`bs="tp"`)
  - Thin plate spline with shrinkage (`bs="ts"`)
  - Duchon spline (`bs="ds", m=c(1, 0.5)`)
  - Thin plate spline with rotated covariates (`bs="tp"`)
- Tensor product smooths
  - Thin plate spline (`bs="tp"`)
  - Thin plate spline with rotated covariates (`bs="tp"`)

In the rotated covariate case, the coordinates were rotated through 45$^o$ by the rotation matrix:
$$
R= \begin{pmatrix}
   \cos(\pi/4) & -\sin(\pi/4)\\
   \sin(\pi/4) & \cos(\pi/4)
   \end{pmatrix}
$$

## Non-spatial models

- Horvitz-Thompson
- Horvitz-Thompson stratified with 2 strata: $x\leq 0.5$ and $x>0.5$.

## True distributions

- "`f`"
- "`lr`"
- "`rl`"
- "`rl_002`"
- "`rl_001`"


## Designs

```{r designs, echo=FALSE, fig.width=4, fig.height=3}
n_segs <- 10

zz <- data.frame(x   = c(seq(0, 0.5, len=n_segs),
                         seq(0.5, 1, len=n_segs)),
                 y   = c(seq(0, 1, len=n_segs),
                         seq(1, 0, len=n_segs)),
                 leg = c(rep("1", n_segs),
                         rep("2", n_segs)))

# many zigzags
mzz <- rbind(zz,zz,zz)

mzz$x <- mzz$x/3
ind <- 1:nrow(zz)
mzz$x[ind+nrow(zz)] <- mzz$x[ind+nrow(zz)]+1/3
mzz$x[ind+2*nrow(zz)] <- mzz$x[ind+2*nrow(zz)]+2/3


mzz$leg <- as.numeric(mzz$leg)
mzz$leg[ind+nrow(zz)] <- mzz$leg[ind+nrow(zz)]+2
mzz$leg[ind+2*nrow(zz)] <- mzz$leg[ind+2*nrow(zz)]+4
mzz$leg <- as.character(mzz$leg)


zzl <- rbind.data.frame(mzz,
                        data.frame(x   = rep(2.8, 10),
                                   y   = seq(0, 1, len=10),
                                   leg = rep(as.character(max(as.numeric(mzz$leg))+1), 10)))

plot(c(0,3),c(0,1), type="n", xlab="x", ylab="y")
for(i in unique(zzl$leg)){
  lines(zzl[,c("x","y")][zzl$leg==i,])
}
```

## Metrics

Where does the truth lie in the distribution of the model?

If we know $N_\text{truth}$ then at what quantile does it lie in the distribution implied by the model, so find $\mathbb{P}[N_\text{truth} \leq \hat{N}]$. Here we assume log-normally distributed $\hat{N}$, so we can use `plnorm` in R to calculate the value we require. This summary statistic gives some idea of both bias and variance.

If the distribution of the statistic is skewed to either end then we can infer under or over estimation of abundance. Flat distribution is good, a "dome" in the middle indicates a conservative estimate (CIs slightly too wide), which we like.

# Results

```{r load-and-plot-results, echo=FALSE, fig.width=12, fig.height=4, warning=FALSE, message=FALSE}
f_names <- c("f.RData", "lr.RData", "rl.RData", "rl_002.RData", "rl_001.RData")
for(i in seq_along(f_names)){
  load(f_names[i])
  big_res$N <- big_res$V1
  big_res$CV <- big_res$V2
  big_res$V1 <- big_res$V2 <- NULL
  big_res$se <- big_res$CV*big_res$N
  big_res$sen <- big_res$se/sqrt(big_res$n)
  
  big_res$sp1 <- 1/big_res$sp1
  big_res$sp2 <- 1/big_res$sp2
  
  f_names[i] <- sub(".RData", "", f_names[i])
  
  reasonable_range <- function(x){
    # 1 and 5 are whiskers
    boxplot.stats(x)$stats[c(1,5)]
  }
  clipped_obs <- function(x, rr, name){
    x <- x[[name]]
    length(x[x<=rr[1] | x>=rr[2]])
  }
  
  p <- ggplot(big_res)+
    geom_histogram(aes(quantile))+
    facet_wrap(~names, nrow=2)+
    ggtitle(paste(f_names[i], "quantile diagnostic")) +
    theme_minimal()
  print(p) 
  
  ## N
  rr <- reasonable_range(big_res$N)
  # build a data.frame to count clipped observation
  clipped_d <- ddply(big_res, .(names), clipped_obs, name="N", rr=rr)
  
  #p <- ggplot(big_res)+
  #  geom_histogram(aes(N))+
  #  facet_wrap(~names, scales="free_x",nrow=2)+
  #  ggtitle(paste(f_names[i], "N"))+
  #  scale_x_continuous(limits = rr) +
  #  geom_text(aes(label=V1, x=rr[1]+(rr[2]-rr[1]), y=50 ), data=clipped_d) +
  #  geom_vline(aes(xintercept=200))
  #print(p) 
  
  # bias boxplot
  big_res$bias <- big_res$N-200
  p <- ggplot(big_res)+
    geom_boxplot(aes(names, bias)) +
    #facet_wrap(~names, scales="free_x",nrow=2)+
    ggtitle(paste(f_names[i], "bias"))+
    coord_cartesian(ylim=c(-200,300)) +
    theme_minimal()
    geom_hline(aes(yintercept=0))
  print(p) 
  
  
  #rr <- reasonable_range(big_res$sen)
  ## build a data.frame to count clipped observation
  #clipped_d <- ddply(big_res, .(names), clipped_obs, name="sen", rr=rr)
  #p <- ggplot(big_res)+
  #  geom_histogram(aes(sen))+
  #  facet_wrap(~names, scales="free_x",nrow=2)+
  #  scale_x_continuous(limits = rr) +
  #  geom_text(aes(label=V1, x=rr[1]+(rr[2]-rr[1]), y=50 ), data=clipped_d) +
  #  ggtitle(paste(f_names[i], "se/sqrt(n)"))
  #print(p) 
  #
  ## smoothing parameters
  #sp_dat <- big_res[!(big_res$names %in% c("HT","HT_strat")),]
  #sp_dat$names <- as.character(sp_dat$names)
  #
  ## mash the sp2s into the sp1 data
  #sp_dat2 <- big_res[big_res$names %in% c("m_xy_te","m_xyr_te"),]
  #sp_dat2$names <- as.character(sp_dat2$names)
  #sp_dat2$names[sp_dat2$names=="m_xy_te"] <- "m_xy_te sp2"
  #sp_dat2$names[sp_dat2$names=="m_xyr_te"] <- "m_xyr_te sp2"
  #sp_dat <- rbind(sp_dat,sp_dat2)
  #sp_dat$names <- as.factor(sp_dat$names)
  #
  #sp_dat$names <- droplevels(sp_dat$names)
  #rr <- reasonable_range(sp_dat$sp1)
  ## build a data.frame to count clipped observation
  #clipped_d <- ddply(sp_dat, .(names), clipped_obs, name="sp1", rr=rr)
  #p <- ggplot(sp_dat)+
  #  geom_histogram(aes(sp1))+
  #  facet_wrap(~names, scales="free_x",nrow=2)+
  #  ggtitle(paste(f_names[i], "1/sp1"))
  #print(p) 
  
}
```


```{r detections, echo=FALSE, fig.width=12, fig.height=4, warning=FALSE, message=FALSE}
f_names <- c("f.RData", "lr.RData", "rl.RData", "rl_002.RData", "rl_001.RData")
ndat <- c()
for(i in seq_along(f_names)){
  load(f_names[i])
  ndat <- rbind(ndat,
                cbind(scenario=sub(".RData", "", f_names[i]),
                      big_res[big_res$name=="HT",])
                )
}
p <- ggplot(ndat)+
  geom_histogram(aes(n), binwidth=2)+
  ggtitle("detections") +
  facet_wrap(~scenario, nrow=1) +
  theme_minimal()
print(p) 
```



